---
title: "EWARSPlus_DHIS2_DataExchange"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{EWARSPlus_DHIS2_DataExchange}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ewarsconnector)
```
## Purpose of this library

The `ewarsconnector` is intended to provide a console based 
library for transferring data between a DHIS2 instance and
an installation of the [EWARS Plus Local app](https://github.com/maquins/ewars_Plus_2023).

## Create the credentials file

In order for for the package to be run on a server in an automated fashion, you will need to create a plain text set of credentials. Please note, that this should only be done on 
a server with disk encryption and never on a public machine. 

The format of the credentials file should be as follows: 

```{js eval = F}
{
  "dhis": {
    "baseurl": "play.dhis2.org/2.36/",
    "username": "admin",
    "password": "district"
  }
}
```

The `baseurl` should be the address of the DHIS2 server. The final forward slash is important and is required. It is suggested that you create a specific user for the `ewarsconnector` with the minimum amount of permissions provided which will be required. 


## Create the EWARS Plus Local configuration file

A YAML based configuration file is used to control the interaction between 
EWARS Plus Local and DHIS2. An example YAML file is provided below. 


```
baseurl: play.dhis2.org/2.36/
previous_years: 2
hierarchy_level: 2
data:
  - population:
      frequency: yearly
      uid: D8Q6nNeQ7i3
  - weekly_hospitalised_cases:
      frequency: weekly
      uid: A0Y0q8g6DHw
  - meantemperature:
      frequency: weekly
      uid: VA05qvanuVs
  - rainsum:
      frequency: weekly
      uid: gsiW9SgolNd
```

The `baseurl` variable should correspond to address of the DHIS2 server, and should
of course be the same as in the credentials file. 

The `previous_years` variable should be an integer, and should correspond to the 
number of previous years (from the current day) that data should be retreived
from the server. 

The `hiearchy_level` should correspond to the level of the organisation unit
hierarchy at which data should be fetch. One corresponds to the root level. 

The `block` should consist of a list of variables which should be fetched from 
DHIS2. These will be obtained from the `api/analytics` end point. The variables names themselves can be anything which you like, but they should be unique. In this case, we have used `population` to represent the corresponding `Variable for annual population` required by EWARS Plus Local. Similarly, the variable `weekly_hospitalized_cases` is used to represent the required `Variable for the weekly number of outbreak` variable in the EWARS Plus Local app. Other alarm indicators such as `rainsum` and `meantemperature` can be provided as alarm indicators. 

For each variable included in the `data` section, you should also provide a frequency at which the data should be obtained (either yearly or weekly). Generally, population data will be yearly, while other variables will be weekly The `uid` should correspond to the DHIS2 data element UID on the server. 

## Fetching data from DHIS2 for use in EWARS Plus Local


In order to exchange data with the DHIS2 server, the `datimutils` package is used. The first step is to login to the server with your credentials. If you have created a credentials file on your server, simply supply this as the argument to the `ewars_login_dhis2` function. If you are working locally, and want to use your keyring, it is also possible. Consult the `datimutils` documentation for specific details about the various authentication mechanisms available with the `loginToDATIM` function. 

```
library(ewarsconnector)
#Establish a connection with the server
ewars_login_dhis2("/home/jason/.secrets/ewars.json")
#Parse the config
ewars_config <- ewars_parse_config()
#Pull the data from the server
ewars_data <- ewars_fetch_data(ewars_config)
writexl::write_xlsx(ewars_data, path ="ewars_data.xlsx")

```

